2.4 PRNT50

### 2024-06-13 2_scripts/2.4 PRNT50.qmd
Update codes using APL v2 (unfinished) Further investigation on vaccination time gap

```{r}
# The CBS Data, extracted using "CITF_CBS_desc_stats.Rmd"
load(file = "../1_data/private/CBSs.RData") # a clean subset of CBS
# Notes: Y.Yuan confirmed on 01-May-2024 the ver 2023-11-30 of CBS is no longer the last ver. The most recent CBS data has 2023 data included but it's not adding new info to our project so we are still using the 2023-11-30 ver of CBS.

# APL Data
load()
load(file = "../1_data/private/APLs.RData") # a clean subset
```

@@

```{r}
library(dplyr)         # Data management
#library(reshape2)      # for melt() function
library(ggplot2)       # ggplot
library(zoo)           # for year-month format
```

@@

#### 1.1 Use the APLs created

```{r}
APLs<-merge(APLs, select(RFD4682_e, order_ID, 
                   'RBD-IgGII_RESULT', `RBD-IgGII_INTERP`,`RBD-IgGII_RESULT_DATE`
                   ), by="order_ID",all=T)
colnames(APLs)<-c(names(APLs)[1:15], "abbott_s", "pos_abbott_s", "month_abbott_s")
APLs$abb_s_index<-grepl("index", APLs$abbott_s, ignore.case = T)
APLs$abb_s_AU<-grepl("ml", APLs$abbott_s, ignore.case = T)
table(APLs$abb_s_index, APLs$abb_s_AU)
# 36 samples missing abbott_s info
APLs$abb_s_index[APLs$abb_s_index==0 & APLs$abb_s_AU==0]<-NA
APLs$abb_s_AU[is.na(APLs$abb_s_index)]<-NA

APLs$abb_s_AU<-as.numeric(APLs$abb_s_AU)
APLs$abb_s_index<-as.numeric(APLs$abb_s_index)
table(APLs$abb_s_index, APLs$abb_s_AU, useNA = 'ifany')

# remove units
APLs$abbott_s<-gsub('[a-z]', '', APLs$abbott_s)
APLs$abbott_s<-gsub('[A-Z]', '', APLs$abbott_s)
APLs$abbott_s<-gsub('[/<>]', '', APLs$abbott_s)
APLs$abbott_s<-as.numeric(APLs$abbott_s)

# check the COI
summary(APLs[APLs$abb_s_index==1 & APLs$pos_abbott_s=="Positive",]$abbott_s)
# 1.4 - 9.6 the COI is 1.4 not 1?
summary(APLs[APLs$abb_s_index==1 & APLs$pos_abbott_s=="Negative",]$abbott_s)
# 8.1 index but negative?

summary(APLs[APLs$abb_s_AU==1 & APLs$pos_abbott_s=="Positive",]$abbott_s)
# 50 - 800000 the COI = 50 good
summary(APLs[APLs$abb_s_AU==1 & APLs$pos_abbott_s=="Negative",]$abbott_s)
# 0-49, COI = 50 good

# COI for anti_n
summary(APLs[APLs$pos_abbott==1 & !is.na(APLs$abbott_n),]$abbott_n)
# 1.4 - 14.19 the COI = 1.4
summary(APLs[APLs$pos_abbott==0 & !is.na(APLs$abbott_n),]$abbott_n)
# 0.001 - 1.39

table(APLs$pos_abbott)
```

```{r}
# abbott_n misclassified as abbott_s with index?
APLs %>%
  .[is.na(.$abbott_n) & .$abb_s_index==1 & !is.na(.$abb_s_index),] # N=104946

APLs %>%
  .[is.na(.$abbott_n) & .$abb_s_AU==1 & !is.na(.$abb_s_AU),] # N=69108

APLs %>%
  .[!is.na(.$abbott_n) & .$abb_s_index==TRUE & !is.na(.$abb_s_index),] # N=0

APLs %>%
  .[!is.na(.$abbott_n) & .$abb_s_AU==1 & !is.na(.$abb_s_AU),]  # N=40690

APLs %>%
  .[!is.na(.$abbott_n) & is.na(.$abb_s_AU),]                   # N=36

table(is.na(APLs[APLs$abb_s_AU==1,]$abbott_n))

# month of data collection for those five groups:
# Group 1: "Abbott-S" (Index) only
fivenum(APLs[APLs$abb_s_index==1,]$samplemonth, na.rm = T)

# Group 2: True Abbott-S (AU) only
fivenum(APLs[is.na(APLs$abbott_n) & APLs$abb_s_AU==1,]$samplemonth, na.rm = T)

# Group 3: True Abbott-S (AU) and Abbott-N?
fivenum(APLs[!is.na(APLs$abbott_n) & APLs$abb_s_AU==1,]$samplemonth, na.rm = T)

# Group 4: no available anti-S data, only Abbott-N
fivenum(APLs[!is.na(APLs$abbott_n) & is.na(APLs$abb_s_AU),]$samplemonth, na.rm = T)
```

**Notes**: So no matter what it really is it can't be Abbott-S-RBD and shall not be included in anti-S analysis. Updates on May 28th: Carmen replied and confirmed they are misplaced anti-N assay results.

#### 1.2 Move the misplaced anti-N back to where they are

```{r}
APLs$abbott_n[APLs$abb_s_index==1 &
              !is.na(APLs$abbott_s)]<-APLs$abbott_s[APLs$abb_s_index==1 & 
                                                   !is.na(APLs$abbott_s)]
APLs$abbott_s[APLs$abb_s_index==1 & !is.na(APLs$abbott_s)]<-NA

# add anti-N interpretation
APLs$pos_abbott[APLs$abbott_n<1.4 & !is.na(APLs$abbott_n)]<-0
APLs$pos_abbott[APLs$abbott_n>=1.4 & !is.na(APLs$abbott_n)]<-1

# redo the anti-S interpretation
APLs$pos_abbott_s<-rep(NA,214780)
APLs$pos_abbott_s[APLs$abbott_s<50 & !is.na(APLs$abbott_s)]<-0
APLs$pos_abbott_s[APLs$abbott_s>=50 & !is.na(APLs$abbott_s)]<-1
table(APLs$pos_abbott_s, useNA = 'ifany')
```

@@

```{r}
APLs$pos_abbott_s2<-rep(NA, 214780)
APLs$pos_abbott_s2[APLs$abb_s_AU==1 & !is.na(APLs$abb_s_AU) &
                   APLs$abbott_s>=402.36 & !is.na(APLs$abbott_s)]<-"Positive"
APLs$pos_abbott_s2[APLs$abb_s_AU==1 & !is.na(APLs$abb_s_AU) &
                     APLs$abbott_s<402.36 & !is.na(APLs$abbott_s)]<-"Negative"
table(APLs$pos_abbott_s2, useNA = 'ifany')

@@

tb_abbs<-APLs %>%
  .[.$abb_s_AU==1,] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbs$prop.abbs1<-tb_abbs$pos_abbott_s/tb_abbs$n
tb_abbs$prop.abbs2<-tb_abbs$pos_abbott_s2/tb_abbs$n

@@

tb_APLCBS2<-rbind.data.frame(tb_abbs_melt, tb_ros_melt)

@@

```{r}
# Add vax variable
APLs<-merge(APLs, select(RFD4682_e, order_ID, VACCINATION_STATUS),
      by="order_ID", all=T)

@@

# Add days since imm (vax)
APLs<-merge(APLs, select(RFD4682_e, order_ID, DAYS_SINCE_IMM),
      by="order_ID", all=T)
```

@@

```{r}
# re-do the melt data: no vax
tb_abbsv0<-APLs %>%
  .[.$abb_s_AU==1 & (.$vax %in% c("Not Immunized", "Pre-Immunization", "1 Dose")),] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsv0$prop.abbs1<-tb_abbsv0$pos_abbott_s/tb_abbsv0$n
tb_abbsv0$prop.abbs2<-tb_abbsv0$pos_abbott_s2/tb_abbsv0$n

# clean NAs
tb_abbsv0<-tb_abbsv0[!is.na(tb_abbsv0$samplemonth),]

@@

# re-do the melt data: at least 1 dose
tb_abbsv1<-APLs %>%
  .[.$abb_s_AU==1 & .$vax %in% c("1-2 Doses","2 Doses","3 Doses",
                             "4 Doses") ,] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsv1$prop.abbs1<-tb_abbsv1$pos_abbott_s/tb_abbsv1$n
tb_abbsv1$prop.abbs2<-tb_abbsv1$pos_abbott_s2/tb_abbsv1$n

@@

tb_abbsv1_melt<-reshape2::melt(select(tb_abbsv1, -c(2:4)), 
                        id.vars = c("samplemonth"),
                        variable.name = "COI",
                        value.name = "prop")
```

@@

```{r}
# Plot: at least 1 dose
p09v1<-tb_abbsv1_melt %>%
  ggplot(aes(x=samplemonth, y=prop, group=COI, colour=COI)) +
  geom_point(colour="red") +
  geom_line(aes(linetype=COI), colour="red")  +
  theme_bw() + 
  ylim(c(0,1)) +
  xlab("Time of Sample Collection") + 
  ylab("Proportion")+ 
  scale_linetype_manual("", values=c(1,3), labels = c("APL-Abbott (COI=50)", 
                                                      "APL-Abbott (COI=402)"))+
  theme(legend.position='top')+
  ggtitle("Stratum: at least 2 doses")
p09v1
```

@@

```{r}
tb_abbsv11<-APLs %>%
  .[.$abb_s_AU==1 & .$vax %in% c("1 Dose") ,] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsv11$prop.abbs1<-tb_abbsv11$pos_abbott_s/tb_abbsv11$n
tb_abbsv11$prop.abbs2<-tb_abbsv11$pos_abbott_s2/tb_abbsv11$n

# clean NAs
tb_abbsv11<-tb_abbsv11[!is.na(tb_abbsv11$samplemonth),]

tb_abbsv11_melt<-reshape2::melt(select(tb_abbsv11, -c(2:4)), 
                        id.vars = c("samplemonth"),
                        variable.name = "COI",
                        value.name = "prop")
```

```{r}
# Plot: 1 dose only
p09v11<-tb_abbsv11_melt %>%
  ggplot(aes(x=samplemonth, y=prop, group=COI, colour=COI)) +
  geom_point(colour="red") +
  geom_line(aes(linetype=COI), colour="red")  +
  theme_bw() + 
  ylim(c(0,1)) +
  xlab("Time of Sample Collection") + 
  ylab("Proportion")+ 
  scale_linetype_manual("", values=c(1,3), labels = c("APL-Abbott (COI=50)", 
                                                      "APL-Abbott (COI=402)"))+
  theme(legend.position='top')+
  ggtitle("Stratum: 1 dose only")
p09v11
```

@@

>>>>

```{r}
# The CBS Data, extracted using "CITF_CBS_desc_stats.Rmd"

load(file = "../1_data/private/CBS.RData") # original CBS
load(file = "../1_data/private/CBSs.RData") # a clean subset of CBS
# Notes: Y.Yuan confirmed on 01-May-2024 the ver 2023-11-30 of CBS is no longer the last ver. The most recent CBS data has 2023 data included but it's not adding new info to our project so we are still using the 2023-11-30 ver of CBS.

# APL Data
load(file = "../1_data/private/APLs.RData") # a clean subset
```

@@

```{r}
library(dplyr)         # Data management
library(reshape2)      # for melt() function
library(ggplot2)       # ggplot
library(zoo)           # for year-month format
```

@@

#### 1.1 Use the APLs created, add anti-S assay results and interp

```{r}
APLs<-merge(APLs, select(RFD4682_ev2, order_ID, 
                   `RBD-IgGII_RESULT`, `RBD-IgGII_INTERP`,`RBD-IgGII_RESULT_DATE`
                   ), by="order_ID",all=T)
colnames(APLs)<-c(names(APLs)[1:10], "abbott_s", "pos_abbott_s", "month_abbott_s")

# check the COI
summary(APLs[APLs$pos_abbott_s=="Positive",]$abbott_s)
# 50 - 800000 the COI = 50 good
summary(APLs[APLs$pos_abbott_s=="Negative",]$abbott_s)
# 0-49, COI = 50 good
```

**Notes**: So no matter what it really is it can't be Abbott-S-RBD and shall not be included in anti-S analysis. Updates on May 28th: Carmen replied and confirmed they are misplaced anti-N assay results. APL data has been cleaned and uploaded to the server to share with Yuan and Matthew.

@@

```{r}
APLs$pos_abbott_s2<-rep(NA, 214776)
APLs$pos_abbott_s2[APLs$abbott_s>=402.36 & !is.na(APLs$abbott_s)]<-"Positive"
APLs$pos_abbott_s2[APLs$abbott_s<402.36 & !is.na(APLs$abbott_s)]<-"Negative"
table(APLs$pos_abbott_s2, useNA = 'ifany')

@@

tb_abbs<-APLs %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbs$prop.abbs1<-tb_abbs$pos_abbott_s/tb_abbs$n
tb_abbs$prop.abbs2<-tb_abbs$pos_abbott_s2/tb_abbs$n

@@

tb_APLCBS2<-rbind.data.frame(tb_abbs_melt, tb_ros_melt)
rm(tb_abbs, tb_abbs_melt, tb_ros, tb_ros_melt)

@@

```{r}
# Add vax variable
APLs<-merge(APLs, select(RFD4682_ev2, order_ID, VACCINATION_STATUS),
      by="order_ID", all=T)

@@

# Add days since imm (vax)
APLs<-merge(APLs, select(RFD4682_ev2, order_ID, DAYS_SINCE_IMM),
      by="order_ID", all=T)
```

@@

```{r}
# re-do the melt data: no vax 
tb_abbsv0<-APLs %>%
  .[.$vax %in% c("Not Immunized", "Pre-Immunization"),] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsv0$prop.abbs1<-tb_abbsv0$pos_abbott_s/tb_abbsv0$n
tb_abbsv0$prop.abbs2<-tb_abbsv0$pos_abbott_s2/tb_abbsv0$n

# clean NAs
tb_abbsv0<-tb_abbsv0[!is.na(tb_abbsv0$samplemonth),]
tb_abbsv0<-tb_abbsv0[!tb_abbsv0$pos_abbott_s==0,]

@@

# re-do the melt data: 1 dose only
tb_abbsv1<-APLs %>%
  .[.$vax %in% c("1 Dose") ,] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsv1$prop.abbs1<-tb_abbsv1$pos_abbott_s/tb_abbsv1$n
tb_abbsv1$prop.abbs2<-tb_abbsv1$pos_abbott_s2/tb_abbsv1$n

@@

tb_abbsv1_melt<-reshape2::melt(select(tb_abbsv1, -c(2:4)), 
                        id.vars = c("samplemonth"),
                        variable.name = "COI",
                        value.name = "prop")

# re-do the melt data: 2 doses +
tb_abbsv2<-APLs %>%
  .[.$vax %in% c("1-2 Doses", "2 Doses", "3 Doses", "4 Doses") ,] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsv2$prop.abbs1<-tb_abbsv2$pos_abbott_s/tb_abbsv2$n
tb_abbsv2$prop.abbs2<-tb_abbsv2$pos_abbott_s2/tb_abbsv2$n

# clean NAs
tb_abbsv2<-tb_abbsv2[!is.na(tb_abbsv2$samplemonth),]

tb_abbsv2_melt<-reshape2::melt(select(tb_abbsv2, -c(2:4)), 
                        id.vars = c("samplemonth"),
                        variable.name = "COI",
                        value.name = "prop")
```

@@

```{r}
# Plot: 1 dose only
p09v1<-tb_abbsv1_melt %>%
  ggplot(aes(x=samplemonth, y=prop, group=COI, colour=COI)) +
  geom_point(colour="red") +
  geom_line(aes(linetype=COI), colour="red")  +
  theme_bw() + 
  ylim(c(0,1)) +
  xlab("Time of Sample Collection") + 
  ylab("Proportion")+ 
  scale_linetype_manual("", values=c(1,3), labels = c("APL-Abbott (COI=50)", 
                                                      "APL-Abbott (COI=402)"))+
  theme(legend.position='top')+
  ggtitle("Stratum: 1 dose only")
p09v1
```

```{r}
# Plot: 2 doses +
p09v2<-tb_abbsv2_melt %>%
  ggplot(aes(x=samplemonth, y=prop, group=COI, colour=COI)) +
  geom_point(colour="red") +
  geom_line(aes(linetype=COI), colour="red")  +
  theme_bw() + 
  ylim(c(0,1)) +
  xlab("Time of Sample Collection") + 
  ylab("Proportion")+ 
  scale_linetype_manual("", values=c(1,3), labels = c("APL-Abbott (COI=50)", 
                                                      "APL-Abbott (COI=402)"))+
  theme(legend.position='top')+
  ggtitle("Stratum: 2 doses +")
p09v2
```

@@

@@

## 4.4 The vaccination policy and those "1-Doser"

#### 4.4.1 add prop (1-dose only)
```{r}
# number of personals in each month
tb_abbsvall<-APLs %>%
  .[!(.$vax %in% c("No_ULI")),] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsvall$n.all<-tb_abbsvall$n

tb_abbsv1<-merge(tb_abbsv1, select(tb_abbsvall, samplemonth, n.all),
      by="samplemonth", all.x = T, all.y = F)
tb_abbsv1$prop.d1<-tb_abbsv1$n/tb_abbsv1$n.all
```

#### 4.4.2 Making the plot:
```{r}
p09v1a<-tb_abbsv1 %>%
  ggplot(aes(x=samplemonth)) +
  geom_point(aes(y=prop.abbs1), colour="red") +
  geom_point(aes(y=prop.abbs2), colour="red") +
  geom_line(aes(y=prop.abbs1), colour="red")  +
  geom_line(aes(y=prop.abbs2), linetype=2, colour="red")  +
  geom_line(aes(y=prop.d1), colour="darkgrey") +
  theme_bw() + 
  ylim(c(0,1)) +
  xlab("Time of Sample Collection") + 
  ylab("Proportion")+ 
  scale_linetype_manual("", values=c(1,3), labels = c("APL-Abbott (COI=50)", 
                                                      "APL-Abbott (COI=402)"))+
  theme(legend.position='top')+
  ggtitle("Stratum: 1 dose only")
p09v1a
```

## 4.4.3 Check the time gap btw 1st dose and 2nd dose. 
```{r}
APL_vax<-RFD4682_v2_vaxe %>%
  select(ULI, ULI_IDe, DOSE_COUNT, IMMUNIZATION_DATE) %>%
  .[.$DOSE_COUNT==1 | .$DOSE_COUNT==2, ]

APL_vax<-reshape(APL_vax[,c(2:4)], idvar="ULI_IDe", timevar = "DOSE_COUNT",
        direction = "wide")

# Calculate time gap btw 1st and 2nd dose
APL_vax$time_gap<-as.Date(APL_vax$IMMUNIZATION_DATE.2)-
                  as.Date(APL_vax$IMMUNIZATION_DATE.1)

# check distribution
hist(as.numeric(APL_vax$time_gap), breaks=25)

summary(as.numeric(APL_vax$time_gap))
```
